Testing atlantisom: generate census length comps

### 2019-05-08 TrueLengthCompTest.Rmd
working through length comp sampling

---
title: "Testing atlantisom: generate census length and age comps"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

@@

## Simulate a survey part 3: sample for length and age composition

@@

To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species.

@@

ageclcomp <- ggplot(dat2totN, aes(x=agecl, y=numAtAge/totN, col=time)) +
  geom_point()

ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
#ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")

@@

```{r comptest1plot, message=FALSE, warning=FALSE}

@@

And they did for NEUS the first time I did this, with some exceptions.

The problems start when I try to do age to length, this function returns an error at present.

**Editing to be done in atlantisom (I have made local changes to calcage2length.R that are not pushed yet because I may have introduced an error)**

```{r calcage2length-breaks, eval=FALSE}
# now turn this into length comp with calc_age2length
# function takes resn and nums as inputs so run survey and sample functions on resn as well

@@

lengthtesthigh <- sample_fish(survey_testNall, effNhigh)
# use fish sampled from N along with survey of resn and structn in function
# calc_age2length uses nums (lengthtest) and finds matches in resn and structn

>>>>

---
title: "Testing atlantisom: generate census length comps"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

@@

## Simulate a survey part 4: sample for length composition

@@

Atlantis outputs numbers by cohort (stage-age) and growth informtation, but does not output size of animals directly. The function we are testing here, `atlantisom::calc_age2length` converts numbers by cohort to a length composition. The [workflow originally envisioned](https://onedrive.live.com/?authkey=%21AFQkOoKRz64TLUw&cid=59547B4CB95EF108&id=59547B4CB95EF108%216291&parId=59547B4CB95EF108%216262&o=OneUp) was to create a survey, sample fish from the survey, then apply this function, but as written it may work better to create a length comp for the whole population, which could then be sampled. We'll work out the options here.

### This stuff is biolerplate from other docs, repeated here for completeness: 
To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species.

@@

# ageclcomp <- ggplot(dat2totN, aes(x=agecl, y=numAtAge/totN, col=time)) +
#   geom_point()
# 
# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
# #ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")

@@

```{r comptest1plot, message=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}

@@

And they mostly do aside from fish that have 0s in the first year (Planktiv_T_Fish.

The problems start when I try to do age to length, this function returns an error as written when `sample_fish` is run first.

May 8 update: Found problem, fix may require rethinking sampling flow. Making another rmd (this one!) to deal with this.

So, test 1 is running `calc_age2length` with the full true dataset. This should work because there will be no aggregation of the output:

```{r calcage2length-truth}

length_census <- calc_age2length(structn = truth$structn,
                                 resn = truth$resn,
                                 nums = truth$nums,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

This produces output. Visualize:

```{r lengthcensus-vis}

# for an example species, by polygon, layer because this is full resolution

So the question now is, do we run `create_survey` on the output of this, or do we `create_survey` on all three truth components first, then run `calc_age2length` to get a survey sample? This should work and will eliminate the unsampled areas/times/species.

And then `sample_fish` from `calc_age2length` output? But also we need to run `sample_fish` to get the cohort (age) output (or do we? see below) so the multinomial in there will make different distributions for smaller sample sizes. Since we are trying to represent biological sampling as one set of sampled fish which we length, then take a subset of ages and weights, this may not work.

Or, we can sum the length atoutput from calc_age2length over size bins and back in to agecl bins to get age classes, then subsample this for ages? We also have mulen and muweight for each size bin (or is it agecl bin?) so can we back the individual weights for the aged fish subsample out of this?

```{r calcage2length-breaks, eval=FALSE}
# now turn this into length comp with calc_age2length
# function takes resn and nums as inputs so run survey and sample functions on resn as well
# do we need to run the survey if calc_age2length does matching based on sample_fish?

@@

lengthtesthighN <- sample_fish(survey_testNall, effNhigh)
# use fish sampled from N along with survey of resn and structn in function
# calc_age2length uses nums (lengthtest) and finds matches in resn and structn

# BUT: expecting raw nums from truth, not sample_fish nums with polygon already aggregated, so this never matches as written.

# solutions, run sample_fish on structn and resn prior to input--*trying*
# or don't sameple_fish on nums first, but I think we want to sample_fish first in the workflow, then convert that sample to lengths, ages, and weights.

# unfortunately, if we are really sampling with low effN, these will be sampling cohorts with different probabilities across nums, structn, and resn, so do we need a length comp on the whole thing first and then sample it??

lengthtesthighstructn <- sample_fish(survey_teststructnall, effNhigh)
lengthtesthighresn <- sample_fish(survey_testresnall, effNhigh)
